### 2019-08-05 reports/thesis/chapters/3_compoundPoisson.tex
compound poisson

The compound Poisson distribution is studied here because of the compound Poission-like behaviour from the detection of photons \citep{whiting2002signal, elbakri2003efficient, whiting2006properties}. It is defined by defining a latent variable $Y\sim\poisson(\lambda)$ with probability mass function (p.m.f.) $\prob(Y=y)=\euler^{-\lambda}\frac{\lambda^y}{y!}$ for $y=0,1,2,\cdots$, where $\lambda>0$ is the Poisson rate parameter. Let $U_i$ be some independent and identically distributed (i.i.d.) latent random variables with probability density function (p.d.f.) $p_U(u)$ for $i=1,2,3,\cdots$. Let $X$ be a compound Poisson random variable where
\begin{equation}
  X|Y = \sum_{i=1}^{Y}U_i \ .
  \label{eq:compoundPoisson_X|Y}
\end{equation}
The p.d.f.~of $X$ can be obtained by marginalising the joint p.d.f.
\begin{equation}
  p_X(x)=\sum_{y=0}^\infty p_{X|Y}(x|y)\prob(Y=y) \quad\text{for }x\geqslant 0
  \ .
\end{equation}
It should be noted that $X=0$ if and only if $Y=0$ with probability $\prob(X=0)=\euler^{-\lambda}$. Then $X$ has probability mass at zero and probability density for positive numbers which results in the p.d.f.
\begin{equation}
  p_X(x) = 
  \begin{cases}
    \delta(x) \euler^{-\lambda}  & \text{ for } x=0 \\ 
    \sum_{y=1}^\infty p_{X|Y}(x|y)\euler^{-\lambda}\frac{\lambda^y}{y!} \quad\text{for } & \text{ for } x>0
  \end{cases}
\end{equation}
where $\delta(x)$ is the Dirac delta function.

This chapter starts with a literature review on the compound Poisson distribution, how it is derived from the behaviour of photons, how its likelihood is evaluated and methods for fitting onto data. A model was proposed for the pixel's grey values and the expectation-maximization (EM) algorithm was implemented to fit the model onto data. It was found that for high photon rate, there were identifiability issues. The chapter is concluded on a discussion on why the EM algorithm failed.

@@

When the photons interact with the scintillator in the detector, they are converted into visible light. The visible light photons are then detected and converted into a digital signal. A quantum counter set the digital signal to be linear with to the number of photons detected \citep{whiting2006properties}. Let $X$ be the grey value observed, then
\begin{equation}
X = bY + \epsilon
\end{equation}
where $\epsilon\sim\normal(a,\kappa)$, $b$ and $a$ are some constant and $\kappa$ is the variance of electronic noise. In a quantum counter, the mean and variance of the grey value are
\begin{equation}
\expectation\left[X\right] = b\lambda + a
\end{equation}
and
\begin{equation}
\variance\left[X\right] = b^2\lambda + \kappa
\end{equation}
respectively. By eliminating $\lambda$
\begin{equation}
\variance\left[X\right] = b\expectation\left[X\right]+\kappa-ab \ ,
\end{equation}
a linear relationship between the variance and expectation of the grey value \citep{ma2012varaince} is obtained.

@@

Experiments have been done to verify the compound Poisson nature of the detector. This was done by investigating the variance of radiographs of air \citep{hsieh2015compound} and a polyethylene cylinder \citep{yang2009evaluation, yang2010noise} at different voltages and powers. It was found there were 2 components in the noise, one was signal dependent and comes from the compound Poisson, the other was signal independent and may be electronic noise. The electronic noise can be modelled as Normally distributed \citep{xu2009electronic}.

@@

Returning back to the compound Poisson with no electronic noise $X|Y = \sum_{i=1}^{Y}U_i$. Let the moment generating function (m.g.f.) of $X$ be $M_X(\theta)=\expectation\left[\euler^{X\theta}\right]$. It can be shown that the m.g.f.~is
\begin{equation}
  M_X(\theta)=
  \exp\left[
    \lambda
    \left(
      M_U(\theta)-1
    \right)
  \right]
\end{equation}
\citep{gatto2010saddlepoint}. The derivation is shown in Appendix \ref{chapter:appendix_compoundPoissonMgf}. Moments of $X$ can be obtained from the m.g.f.~by differentiating it and setting it to zero. In other words, $\expectation[X^r]=M_X^{(r)}(\theta)$. Then it can be shown that
\begin{align}
  \expectation\left[X\right]&=\lambda\expectation\left[U\right]
  \\
  \variance\left[X\right] &= \lambda\expectation\left[U^2\right]
  \\
  \expectation\left[(X-\expectation[X])^3\right] &=\lambda\expectation\left[U^3\right]
\end{align}
which agrees with the expectation and variance results in the previous section. The derivation is shown in Appendix \ref{chapter:appendix_compoundPoissonMgf}.

@@

The m.g.f.~of $U$ is $M_U(\theta) = \left(\dfrac{\beta}{\beta-\theta}\right)^\alpha$. Then the m.g.f.~of $X$ is
\begin{equation}
  M_X(\theta)=\exp\left[\lambda\left(\left(\frac{\beta}{\beta-\theta}\right)^{\alpha}-1\right)\right]
\end{equation}
and moments can be obtained from it such as
\begin{equation}
  \expectation\left[X\right]=\frac{\alpha\lambda}{\beta}
\end{equation}
\begin{equation}
  \variance\left[X\right]=\frac{\alpha(\alpha+1)\lambda}{\beta^2}
  \label{eq:compoundPoisson_variance}
\end{equation}
and
\begin{equation}
  \expectation\left[(X-\expectation[X])^3\right] = \frac{\alpha(\alpha+1)(\alpha+2)\lambda}{\beta^3}
  \ .
\end{equation}

@@

The infinite sum, $\sum_{y=1}^\infty W_y$, can be computationally summed in a clever way in order to evaluate the p.d.f. This was done by summing only large terms in the sum and ignore small terms \citep{dunn2005series}. \cite{dunn2005series} approximated the sum by truncation
\begin{equation}
  \sum_{y=1}^\infty W_y \approx \sum_{y=y_\text{l}}^{y_\text{u}}W_y
\end{equation}
where $y_\text{l}<y_{\text{max}}<y_\text{u}$ and $y_{\text{max}}$ is the value of $y$ which maximises $W_y$. \cite{dunn2005series} used Stirling's approximation to find that
\begin{equation}
  y_{\text{max}} \approx \frac{x^{2-p}}{\phi(2-p)}
\end{equation}
by treating $W_y$ as a continuous and differentiable function of $y$. The derivation is shown in \ref{chapter:appendix_compoundPoissonSeries}. Because values of $y$ should be a positive integer, it would be appropriate to round the solution to $y_\text{max}$ accordingly
\begin{equation}
  y_{\text{max}} = \text{max}\left[
    1,\text{round}\left(\frac{x^{2-p}}{\phi(2-p)}\right)
  \right]
  \ .
\end{equation}

@@

The evaluation of the p.d.f.~using the saddlepoint approximate required a bit of caution to avoid over/underflow problems. Suppose realizations of $X$ were simulated $\left\{x_1,x_2,x_3,\cdots,x_n\right\}$. The saddlepoint approximation was computed up to a constant using
\begin{equation}
  p_X(x) \propto
  \exp\left[
    -\frac{\alpha+2}{2(\alpha+1)}
    \ln(x)
    -x\beta
    +\left(
    \frac{x\beta}{\alpha}
    \right)^{\frac{\alpha}{\alpha+1}}\lambda^{\frac{1}{\alpha+1}}(\alpha+1) - k
  \right]
\end{equation}
for 500 equally spaced points from and including the minimum non-zero simulated value to the maximum simulated value. $k$ is some constant was chosen to be
\begin{equation}
  k =
  \max_{i\in\left\{1,2,3,\cdots,n\right\}}
  \left[
    -\frac{\alpha+2}{2(\alpha+1)}
    \ln(x_i)
    -x_i\beta
    +\left(
      \frac{x_i\beta}{\alpha}
    \right)^{\frac{\alpha}{\alpha+1}}\lambda^{\frac{1}{\alpha+1}}(\alpha+1)
  \right]
  \ .
\end{equation}
The density was then was normalised by numerically integrating it using the trapezium rule using the 500 evaluated points.

A Q-Q plot is a plot which compares the empirical quantiles with the theoretical quantiles. Let
\begin{equation}
  F_X(x) = \prob(X\leqslant x)
\end{equation}
and
\begin{equation}
  \widehat{F}_X(x) = \frac{1}{n}\cdot\text{max}
  \left[
    \left(\sum_{i=1}^n\mathbb{I}(x_i\leqslant x)\right)-0.5,0
  \right]
  \ ,
\end{equation}
then a Q-Q plot is a parametric plot which plots $\widehat{F}_X^{-1}(p)$ against $F_X^{-1}(p)$ for $p=\frac{0.5}{n},\frac{1.5}{n},\frac{2.5}{n},\cdots,\frac{n-0.5}{n}$. If $F_X(x)$ and $\widehat{F}_X(x)$ are similar, a Q-Q plot should be a straight line with gradient 1, intercepting the origin. For the exact method and the saddlepoint approximate, $F_X(x)$ was found numerically by evaluating the p.d.f.~at $10\,000$ equally spaced points from and including the minimum to the maximum of the simulated samples, and then summing the required trapeziums. $\widehat{F}_X^{-1}(p)$ was then calculated by interpolation. Because the Normal and saddlepoint approximation does not support zero, the Q-Q plot was omitted for these approximations for small $\lambda$.

@@

For low $\lambda$ (Figures \ref{fig:compoundPoisson_histogram_1_1_1}, \ref{fig:compoundPoisson_histogram_1_100_1}, \ref{fig:compoundPoisson_histogram_approx_1_1_1} and \ref{fig:compoundPoisson_histogram_approx_1_100_1}) there is a chance of simulating zeros. For $\lambda = 1$, the Normal approximation failed to capture the probability mass at zero because the Normal distribution is symmetric and support negative values. The saddlepoint approximation did capture the probability mass at zero as shown by an increase in probability density towards zero. For $\CPoisson(1,100,1)$ in Figure \ref{fig:compoundPoisson_histogram_approx_1_100_1}, the compound Poisson-gamma p.d.f.~contains multiple peaks and the saddlepoint approximation was not flexible enough them. In Figure \ref{fig:compoundPoisson_histogram_1_100_1}, the Q-Q plot for the exact method was quite sensitive at the tails of each peak, perhaps there were a few inaccuracies in the exact evaluation of the p.d.f.~or in the calculation or inversion of the c.d.f.

@@

\begin{figure}
  \centering
    \centerline{
    \begin{subfigure}[b]{\mainSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_CompoundPoissonlambda1alpha1beta1.eps}
        \caption{Left: Observed and expected frequency of a zero. Right: Histogram and p.d.f.~of non-zero values.}
    \end{subfigure}
    }
    \vspace{2em}
    \centerline{
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_qq_CompoundPoissonlambda1alpha1beta1.eps}
        \caption{Q-Q plot}
    \end{subfigure}
    }
    \caption{1\,000 simulations of a $\CPoisson(1,1,1)$ random variable were simulated and its empirical distribution is compared to the p.d.f.~evaluated using the exact method. In a), the dotted red line shows the $\Phi(\pm1)$ quantile of the expected frequency or frequency density.}
    \label{fig:compoundPoisson_histogram_1_1_1}
\end{figure}

\begin{figure}
  \centering
    \centerline{
    \begin{subfigure}[b]{\mainSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_CompoundPoissonlambda1alpha100beta1.eps}
        \caption{Left: Observed and expected frequency of a zero. Right: Histogram and p.d.f.~of non-zero values.}
    \end{subfigure}
    }
    \vspace{2em}
    \centerline{
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_qq_CompoundPoissonlambda1alpha100beta1.eps}
        \caption{Q-Q plot}
    \end{subfigure}
    }
    \caption{1\,000 simulations of a $\CPoisson(1,100,1)$ random variable were simulated and its empirical distribution is compared to the p.d.f.~evaluated using the exact method. In a), the dotted red line shows the $\Phi(\pm1)$ quantile of the expected frequency or frequency density.}
    \label{fig:compoundPoisson_histogram_1_100_1}
\end{figure}

\begin{figure}
  \centering
    \centerline{
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_CompoundPoissonNormlambda1alpha1beta1.eps}
        \caption{Normal approximation}
    \end{subfigure}
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_CompoundPoissonSaddlelambda1alpha1beta1.eps}
        \caption{Saddlepoint approximation}
    \end{subfigure}
    }
    \caption{1\,000 simulations of a $\CPoisson(1,1,1)$ random variable were simulated and its empirical distribution is compared to the p.d.f.~evaluated using approximate methods. In a), the dotted red line shows the $\Phi(\pm1)$ quantile of the expected frequency density.}
    \label{fig:compoundPoisson_histogram_approx_1_1_1}
\end{figure}

\begin{figure}
  \centering
    \centerline{
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_CompoundPoissonNormlambda1alpha100beta1.eps}
        \caption{Normal approximation}
    \end{subfigure}
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_CompoundPoissonSaddlelambda1alpha100beta1.eps}
        \caption{Saddlepoint approximation}
    \end{subfigure}
    }
    \caption{1\,000 simulations of a $\CPoisson(1,100,1)$ random variable were simulated and its empirical distribution is compared to the p.d.f.~evaluated using approximate methods. In a), the dotted red line shows the $\Phi(\pm1)$ quantile of the expected frequency density.}
    \label{fig:compoundPoisson_histogram_approx_1_100_1}
\end{figure}

\begin{figure}
  \centering
    \centerline{
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_CompoundPoissonlambda10alpha1beta1.eps}
        \caption{Histogram - Exact method}
    \end{subfigure}
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_qq_CompoundPoissonlambda10alpha1beta1.eps}
        \caption{Q-Q plot - Exact method}
    \end{subfigure}
    }
    \centerline{
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_CompoundPoissonNormlambda10alpha1beta1.eps}
        \caption{Histogram - Normal approx.}
    \end{subfigure}
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_qq_CompoundPoissonNormlambda10alpha1beta1.eps}
        \caption{Q-Q plot - Normal approx.}
    \end{subfigure}
    }
    \centerline{
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_CompoundPoissonSaddlelambda10alpha1beta1.eps}
        \caption{Histogram - Saddlepoint approx.}
    \end{subfigure}
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_qq_CompoundPoissonSaddlelambda10alpha1beta1.eps}
        \caption{Q-Q plot - Saddlepoint approx.}
    \end{subfigure}
    }
    \caption{1\,000 simulations of a $\CPoisson(10,1,1)$ random variable were simulated and its empirical distribution is compared to the p.d.f. On the left, the dotted red line shows the $\Phi(\pm1)$ quantile of the expected frequency density.}
    \label{fig:compoundPoisson_histogram_10_1_1}
\end{figure}

>>>>

The compound Poisson distribution is studied here because of the compound Poission-like behaviour from the detection of photons \citep{whiting2002signal, elbakri2003efficient, whiting2006properties}. It is defined by defining a latent variable $Y\sim\poisson(\lambda)$ with probability mass function (p.m.f.) $\prob(Y=y)=\euler^{-\lambda}\frac{\lambda^y}{y!}$ for $y=0,1,2,\cdots$, where $\lambda>0$ is the Poisson rate parameter. Let $U_i$ be some independent and identically distributed (i.i.d.) latent random variables with probability density function (p.d.f.) $p_U(u)$ for $i=1,2,3,\cdots$. Let $X$ be a compound Poisson random variable where
\begin{equation}
  X|Y = \sum_{i=1}^{Y}U_i \ .
  \label{eq:compoundPoisson_X|Y}
\end{equation}
The p.d.f.~of $X$ can be obtained by marginalising the joint p.d.f.
\begin{equation}
  p_X(x)=\sum_{y=0}^\infty p_{X|Y}(x|y)\prob(Y=y) \quad\text{for }x\geqslant 0
  \ .
\end{equation}
It should be noted that $X=0$ if and only if $Y=0$ and this happens with probability $\prob(Y=0)=\euler^{-\lambda}$. This implies that $X$ has probability mass at zero and probability density for positive numbers which results in the p.d.f.
\begin{equation}
  p_X(x) = 
  \begin{cases}
    \delta(x) \euler^{-\lambda}  & \text{ for } x=0 \\ 
    \sum_{y=1}^\infty p_{X|Y}(x|y)\euler^{-\lambda}\frac{\lambda^y}{y!} \quad\text{for } & \text{ for } x>0
  \end{cases}
\end{equation}
where $\delta(x)$ is the Dirac delta function.

This chapter starts with a literature review on the compound Poisson distribution, how it is derived from the behaviour of photons, how its likelihood is evaluated and methods for fitting it onto data. A model was proposed for the pixel's grey values and the expectation-maximization (EM) algorithm was implemented to fit the model onto data. It was found that for high photon rate, there were identifiability issues. The chapter is concluded on a discussion on why the EM algorithm failed.

@@

When the photons interact with the scintillator in the detector, they are converted into visible light. The visible light photons are then detected and converted into a digital signal. A quantum counter set the digital signal to be linear with to the number of photons detected \citep{whiting2006properties}. Let $X$ be the grey value observed, then
\begin{equation}
X = bY + \epsilon
\end{equation}
where $\epsilon\sim\normal(a,\kappa)$, $b$ and $a$ are some constant and $\kappa$ is the variance of electronic noise. The mean and variance of the grey value are
\begin{equation}
\expectation\left[X\right] = b\lambda + a
\end{equation}
and
\begin{equation}
\variance\left[X\right] = b^2\lambda + \kappa
\end{equation}
respectively. By eliminating $\lambda$
\begin{equation}
\variance\left[X\right] = b\expectation\left[X\right]+\kappa-ab \ ,
\end{equation}
a linear relationship between the variance and expectation of the grey value \citep{ma2012varaince} is obtained.

@@

Experiments have been done to verify the compound Poisson nature of the detector. This was done by investigating the variance of radiographs of air \citep{hsieh2015compound} and a polyethylene cylinder \citep{yang2009evaluation, yang2010noise} at different voltages and powers. It was found there were two components in the noise, one was signal dependent and comes from the compound Poisson, the other was signal independent and may be electronic noise. The electronic noise can be modelled using a Normal distribution \citep{xu2009electronic}.

@@

Returning back to the compound Poisson with no electronic noise $X|Y = \sum_{i=1}^{Y}U_i$. Let the moment generating function (m.g.f.) of $X$ be $M_X(\theta)=\expectation\left[\euler^{X\theta}\right]$. It can be shown that the m.g.f.~is
\begin{equation}
  M_X(\theta)=
  \exp\left[
    \lambda
    \left(
      M_U(\theta)-1
    \right)
  \right]
\end{equation}
\citep{gatto2010saddlepoint}. The derivation is shown in Appendix \ref{chapter:appendix_compoundPoissonMgf}. Moments of $X$ can be obtained from the m.g.f.~by differentiating it and setting $\theta$ to zero. In other words, $\expectation[X^r]=M_X^{(r)}(0)$. Then it can be shown that
\begin{align}
  \expectation\left[X\right]&=\lambda\expectation\left[U\right]
  \\
  \variance\left[X\right] &= \lambda\expectation\left[U^2\right]
  \\
  \expectation\left[(X-\expectation[X])^3\right] &=\lambda\expectation\left[U^3\right]
\end{align}
which agrees with the expectation and variance results in the previous section. The derivation is shown in Appendix \ref{chapter:appendix_compoundPoissonMgf}.

@@

The m.g.f.~of $U$ is $M_U(\theta) = \left(\dfrac{\beta}{\beta-\theta}\right)^\alpha$, then the m.g.f.~of $X$ is
\begin{equation}
  M_X(\theta)=\exp\left[\lambda\left(\left(\frac{\beta}{\beta-\theta}\right)^{\alpha}-1\right)\right]
\end{equation}
and moments can be obtained from it such as
\begin{equation}
  \expectation\left[X\right]=\frac{\alpha\lambda}{\beta}
\end{equation}
\begin{equation}
  \variance\left[X\right]=\frac{\alpha(\alpha+1)\lambda}{\beta^2}
  \label{eq:compoundPoisson_variance}
\end{equation}
and
\begin{equation}
  \expectation\left[(X-\expectation[X])^3\right] = \frac{\alpha(\alpha+1)(\alpha+2)\lambda}{\beta^3}
  \ .
\end{equation}

@@

The infinite sum, $\sum_{y=1}^\infty W_y$, can be computationally summed in a clever way in order to evaluate the p.d.f. This was done by summing only large terms in the sum and ignore small terms \citep{dunn2005series}. \cite{dunn2005series} approximated the sum by truncation
\begin{equation}
  \sum_{y=1}^\infty W_y \approx \sum_{y=y_\text{l}}^{y_\text{u}}W_y
\end{equation}
where $y_\text{l}<y_{\text{max}}<y_\text{u}$ and $y_{\text{max}}$ is the value of $y$ which maximises $W_y$. \cite{dunn2005series} used Stirling's approximation to find that
\begin{equation}
  y_{\text{max}} \approx \frac{x^{2-p}}{\phi(2-p)}
\end{equation}
by treating $W_y$ as a continuous and differentiable function of $y$. The derivation is shown in Appendix \ref{chapter:appendix_compoundPoissonSeries}. Because values of $y$ should be a positive integer, it would be appropriate to round the solution to $y_\text{max}$ accordingly
\begin{equation}
  y_{\text{max}} = \text{max}\left[
    1,\text{round}\left(\frac{x^{2-p}}{\phi(2-p)}\right)
  \right]
  \ .
\end{equation}

@@

The evaluation of the p.d.f.~using the saddlepoint approximate required a bit of caution to avoid over/underflow problems. Suppose realizations of $X$ were simulated $\left\{x_1,x_2,x_3,\cdots,x_n\right\}$. The saddlepoint approximation was computed up to a constant using
\begin{equation}
  p_X(x) \propto
  \exp\left[
    -\frac{\alpha+2}{2(\alpha+1)}
    \ln(x)
    -x\beta
    +\left(
    \frac{x\beta}{\alpha}
    \right)^{\frac{\alpha}{\alpha+1}}\lambda^{\frac{1}{\alpha+1}}(\alpha+1) - k
  \right]
\end{equation}
for 500 equally spaced points from and including the minimum non-zero simulated value to the maximum simulated value. $k$ is some constant which was chosen to be
\begin{equation}
  k =
  \max_{i\in\left\{1,2,3,\cdots,n\right\}}
  \left[
    -\frac{\alpha+2}{2(\alpha+1)}
    \ln(x_i)
    -x_i\beta
    +\left(
      \frac{x_i\beta}{\alpha}
    \right)^{\frac{\alpha}{\alpha+1}}\lambda^{\frac{1}{\alpha+1}}(\alpha+1)
  \right]
  \ .
\end{equation}
The density was then was normalised by numerically integrating it using the trapezium rule using the 500 evaluated points.

A Q-Q plot is a plot which compares the empirical quantiles with the theoretical quantiles. Let
\begin{equation}
  F_X(x) = \prob(X\leqslant x)
\end{equation}
and
\begin{equation}
  \widehat{F}_X(x) = \frac{1}{n}\cdot\text{max}
  \left[
    \left(\sum_{i=1}^n\mathbb{I}(x_i\leqslant x)\right)-0.5,0
  \right]
  \ ,
\end{equation}
then a Q-Q plot is a parametric plot which plots $\widehat{F}_X^{-1}(p)$ against $F_X^{-1}(p)$ for $p=\frac{0.5}{n},\frac{1.5}{n},\frac{2.5}{n},\cdots,\frac{n-0.5}{n}$. If $F_X(x)$ and $\widehat{F}_X(x)$ are similar, a Q-Q plot should be a straight line with gradient 1, intercepting the origin. For the exact method and the saddlepoint approximate, $F_X(x)$ was found numerically by evaluating the p.d.f.~at $10\,000$ equally spaced points from and including the minimum to the maximum of the simulated samples, and then summing the required trapeziums. $\widehat{F}_X^{-1}(p)$ was then calculated by interpolation. Because the Normal and saddlepoint approximation does not support zero, the Q-Q plot was omitted for small $\lambda$.

@@

For low $\lambda$ (Figures \ref{fig:compoundPoisson_histogram_1_1_1}, \ref{fig:compoundPoisson_histogram_1_100_1} and \ref{fig:compoundPoisson_histogram_approx}) there is a chance of simulating zeros. For $\lambda = 1$, the Normal approximation failed to capture the probability mass at zero because the Normal distribution is symmetric and support negative values. The saddlepoint approximation did capture the probability mass at zero as shown by an increase in probability density towards zero. For $\CPoisson(1,100,1)$ in Figure \ref{fig:compoundPoisson_histogram_approx}, the compound Poisson-gamma p.d.f.~contains multiple peaks and the saddlepoint approximation was not flexible enough to capture them. In Figure \ref{fig:compoundPoisson_histogram_1_100_1}, the Q-Q plot for the exact method was quite sensitive at the tails of each peak, perhaps there were a few inaccuracies in the exact evaluation of the p.d.f.~or in the calculation or inversion of the c.d.f.

@@

\begin{figure}
  \centering
    \centerline{
    \begin{subfigure}[b]{\mainSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_CompoundPoissonlambda1alpha1beta1.eps}
        \caption{Left: Observed and expected frequency of a zero. Right: Histogram and p.d.f.~of non-zero values.}
    \end{subfigure}
    }
    \vspace{2em}
    \centerline{
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_qq_CompoundPoissonlambda1alpha1beta1.eps}
        \caption{Q-Q plot}
    \end{subfigure}
    }
    \caption{1\,000 $\CPoisson(1,1,1)$ random variables were simulated and its empirical distribution is compared to the p.d.f.~evaluated using the exact method. In a), the dotted red line shows the $\Phi(\pm1)$ quantile of the expected frequency or frequency density.}
    \label{fig:compoundPoisson_histogram_1_1_1}
\end{figure}

\begin{figure}
  \centering
    \centerline{
    \begin{subfigure}[b]{\mainSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_CompoundPoissonlambda1alpha100beta1.eps}
        \caption{Left: Observed and expected frequency of a zero. Right: Histogram and p.d.f.~of non-zero values.}
    \end{subfigure}
    }
    \vspace{2em}
    \centerline{
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_qq_CompoundPoissonlambda1alpha100beta1.eps}
        \caption{Q-Q plot}
    \end{subfigure}
    }
    \caption{1\,000 $\CPoisson(1,100,1)$ random variables were simulated and its empirical distribution is compared to the p.d.f.~evaluated using the exact method. In a), the dotted red line shows the $\Phi(\pm1)$ quantile of the expected frequency or frequency density.}
    \label{fig:compoundPoisson_histogram_1_100_1}
\end{figure}

\begin{figure}
  \centering
    \centerline{
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_CompoundPoissonNormlambda1alpha1beta1.eps}
        \caption{$\CPoisson(1,1,1)$ Normal approx.}
    \end{subfigure}
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_CompoundPoissonSaddlelambda1alpha1beta1.eps}
        \caption{$\CPoisson(1,1,1)$ saddlepoint approx.}
    \end{subfigure}
    }
    \vspace{1em}
    \centerline{
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_CompoundPoissonNormlambda1alpha100beta1.eps}
        \caption{$\CPoisson(1,100,1)$ Normal approx.}
    \end{subfigure}
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_CompoundPoissonSaddlelambda1alpha100beta1.eps}
        \caption{$\CPoisson(1,100,1)$ saddlepoint approx.}
        \label{fig:compoundPoisson_histogram_approx_1_100_1}
    \end{subfigure}
    }
    \caption{1\,000 $\CPoisson(1,\alpha,1)$ random variables were simulated and its empirical distribution is compared to the p.d.f.~evaluated using approximate methods. The dotted red line shows the $\Phi(\pm1)$ quantile of the expected frequency density.}
    \label{fig:compoundPoisson_histogram_approx}
\end{figure}

\begin{figure}
  \centering
    \centerline{
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_CompoundPoissonlambda10alpha1beta1.eps}
        \caption{Histogram - Exact method}
    \end{subfigure}
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_qq_CompoundPoissonlambda10alpha1beta1.eps}
        \caption{Q-Q plot - Exact method}
    \end{subfigure}
    }
    \centerline{
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_CompoundPoissonNormlambda10alpha1beta1.eps}
        \caption{Histogram - Normal approx.}
    \end{subfigure}
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_qq_CompoundPoissonNormlambda10alpha1beta1.eps}
        \caption{Q-Q plot - Normal approx.}
    \end{subfigure}
    }
    \centerline{
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_CompoundPoissonSaddlelambda10alpha1beta1.eps}
        \caption{Histogram - Saddlepoint approx.}
    \end{subfigure}
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_qq_CompoundPoissonSaddlelambda10alpha1beta1.eps}
        \caption{Q-Q plot - Saddlepoint approx.}
    \end{subfigure}
    }
    \caption{1\,000 $\CPoisson(10,1,1)$ random variables were simulated and its empirical distribution is compared to the p.d.f. On the left, the dotted red line shows the $\Phi(\pm1)$ quantile of the expected frequency density.}
    \label{fig:compoundPoisson_histogram_10_1_1}
\end{figure}

\begin{figure}
  \centering
    \centerline{
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_CompoundPoissonlambda100alpha100beta1.eps}
        \caption{Histogram - Exact method}
    \end{subfigure}
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_qq_CompoundPoissonlambda100alpha100beta1.eps}
        \caption{Q-Q plot - Exact method}
    \end{subfigure}
    }
    \centerline{
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_CompoundPoissonNormlambda100alpha100beta1.eps}
        \caption{Histogram - Normal approx.}
    \end{subfigure}
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_qq_CompoundPoissonNormlambda100alpha100beta1.eps}
        \caption{Q-Q plot - Normal approx.}
    \end{subfigure}
    }
    \centerline{
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_CompoundPoissonSaddlelambda100alpha100beta1.eps}
        \caption{Histogram - Saddlepoint approx.}
    \end{subfigure}
    \begin{subfigure}[b]{\subSize}
        \includegraphics[width=\textwidth]{../figures/compoundpoisson/cpHistogram_qq_CompoundPoissonSaddlelambda100alpha100beta1.eps}
        \caption{Q-Q plot - Saddlepoint approx.}
    \end{subfigure}
    }
    \caption{1\,000 $\CPoisson(100,100,1)$ random variables were simulated and its empirical distribution is compared to the p.d.f. On the left, the dotted red line shows the $\Phi(\pm1)$ quantile of the expected frequency density.}
    \label{fig:compoundPoisson_histogram_100_1_1}
\end{figure}
